\huge{ \emph{Supplement to} Algorithms For Shaping a Particle Swarm\\ With a Shared Control Input Using Boundary Interaction}

### 2016-12-08 RA-L/theory.tex
Newest version of the algorithm(I found a bug in it.)
workspace, not workplace

\subsection{Using Boundaries: stable configurations of a swarm}\label{subsec:FluidInTank}
One method to control a swarm's shape in a bounded workspace is to simply push in a given direction until the swarm conforms to the boundary. Like fluid settling in a tank, the stable final configuration minimizes potential energy.
\paragraph{Square workplace}
%The formulas for means $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ are as follows: 
We first examine the mean $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ of a very large swarm of robots as they move inside a square workplace under the influence of gravity pointing in the direction $\beta$. The swarm is large, but the robots are small in comparison, and together occupy a constant area $A$ ranging from [0,1]. Under a global input such as gravity, they flow like water, moving to a side of the workplace and forming a polygonal shape, as shown in Fig.~\ref{fig:SquareFill},  see also \citep{Zhao2016mathematicaSquare}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Circular workplace}
Though rectangular boundaries are common in artificial workspaces, biological workspaces are usually rounded.
Similar calculations can be made for a circular workspace.  The workspace is a circle centered at (0,0) with radius 1 and thus area $\pi$.
For notational simplicity, the swarm is parameterized by the global control input signal $\beta$ and the fill-level $h$.  
Under a global input, the robot swarm fills the region under a chord with area
\begin{align}
A(h) = \arccos(1-h)-(1-h) \sqrt{(2-h) h}.
\end{align}
For a circular workspace, the locus of mean positions are aligned with $\beta$ and the mean position is at radius $r(h)$ from the center:
\begin{align}
r(h) = \frac{2 (-(h-2) h)^{3/2}}{3 \left(\sqrt{-(h-2) h} (h-1)+\arccos(1-h)\right)}
\end{align}
Variance $\sigma^2_x(\beta,h)$ is maximized at $\beta = \pi/2+n \pi$ and $h\approx1.43$, while covariance is maximized at $\beta = \pi3/4+n \pi$ and $h\approx0.92.$ For small $h$ values, correlation approaches $\pm1$. Results are displayed in Fig.~\ref{fig:CircleFill}, see also \citep{Zhao2016mathematica}.

@@

\begin{algorithm}
\caption{GenerateDesired spacing($s_1,s_2,e_1,e_2,L$)}\label{alg:PosControl2Robots}
\begin{algorithmic}[1]
\scriptsize
\Require knowledge of starting $(s_1,s_2)$ and ending $(e_1,e_2)$ positions of  two robots. 
$(0,0)$ is bottom corner, $s_1$ is leftmost robot, 
 $L$ is length of the walls. Current robot positions are $(r_1,r_2)$.
\Ensure $(e_1, e_2) , (s_1, s_2)$ all at least $\epsilon$ distance from walls.
\Ensure $s_1.x < s_2.x$ and $s_1.y \geq s_2.y$ if not, rotate workspace coordinates $90^{\circ}$.
\State $\epsilon \gets$ robot diameter
\State $\Delta e_y = e_2.y - e_1.y$
\State $\Delta e_x = e_2.x - e_1.x$
%\If {$\Delta e_y \neq 0$}
%\State $sign \gets \frac{|\Delta e_y|}{\Delta e_y}$
%\Else
%\State $sign \gets 0$
%\EndIf
\State Move $\left(-r_1.x , -\min \left(r_2.y , r_1.y+ \Delta e_y \right)+\epsilon \right)$ \Comment{Touch left wall}
\State Move $(-\min(|-r_2.x+ \Delta e_x|, r_2.x), \sign(\Delta e_y) \, \min(\max(r_1.y + \Delta e_y, r_2.y)-\epsilon,L-r_2.y-\epsilon)+ \epsilon)$ \Comment{Adjust  $y$}
\If {$r_2.y \neq r_1.y$}
\State Move $(\max(\epsilon, -\Delta e_x+ (r_2.x - r_1.x)), -\min(r_1.y , r_2.y))$ \Comment{Touch bottom wall}
\State Move $(\Delta e_x , (|\sign(\Delta e_y)|-1) \epsilon)$ \Comment{Adjust $x$}
\EndIf
\If {$\Delta e_y \neq r_2.y - r_1.y$}
\State Move $(L- r_1.x , \epsilon)$ \Comment{Touch right wall}
\State Move $(0, \Delta e_y - r_2.y)$ \Comment{Adjust $y$}
\EndIf
\State Move by $(e_2.x - r_2.x, e_2.y - r_2.y)$ \Comment{Go to goal}
\end{algorithmic}
\end{algorithm}

>>>>

\subsection{Using Boundaries: stable configurations of a swarm}\label{subsec:FluidInTank}
One method to control a swarm's shape in a bounded workspace is to simply push in a given direction until the swarm conforms to the boundary. Like fluid settling in a tank, the stable final configuration minimizes potential energy.
\paragraph{Square workspace}
%The formulas for means $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ are as follows: 
We first examine the mean $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ of a large swarm of robots as they move inside a square workspace under the influence of a force pulling in the direction $\beta$. The swarm is large, but the robots are small in comparison, and together occupy a constant area $A$ ranging from [0,1]. Under a global input, they move to a side of the workspace and form a polygonal shape, as shown in Fig.~\ref{fig:SquareFill},  see also \citep{Zhao2016mathematicaSquare}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Circular workspace}
Though rectangular boundaries are common in artificial workspaces, biological workspaces are usually rounded.
Similar calculations can be made for a circular workspace.  The workspace is a circle centered at (0,0) with radius 1 and thus area $\pi$.
For notational simplicity, the swarm is parameterized by the global control input signal $\beta$ and the fill-level $h$.  
Under a global input, the robot swarm fills the region under a chord with area
\begin{align}
A(h) = \arccos(1-h)-(1-h) \sqrt{(2-h) h}.
\end{align}
For a circular workspace, the locus of mean positions are aligned with $\beta$ and the mean position is at radius $r(h)$ from the center:
\begin{align}
r(h) = \frac{2 (-(h-2) h)^{3/2}}{3 \left(\sqrt{-(h-2) h} (h-1)+\arccos(1-h)\right)}
\end{align}
Variance $\sigma^2_x(\beta,h)$ is maximized at $\beta = \pi/2+n \pi$ and $h\approx1.43$, while covariance is maximized at $\beta = \pi3/4+n \pi$ and $h\approx0.92.$ For small $h$ values, correlation approaches $\pm1$. Results are displayed in Fig.~\ref{fig:CircleFill}, see also \citep{Zhao2016mathematica}.

@@

\begin{algorithm}
\caption{GenerateDesired spacing($s_1,s_2,e_1,e_2,L$)}\label{alg:PosControl2Robots}
\begin{algorithmic}[1]
\scriptsize
\Require knowledge of starting $(s_1,s_2)$ and ending $(e_1,e_2)$ positions of  two robots. 
$(0,0)$ is bottom corner, $s_1$ is leftmost robot, 
 $L$ is length of the walls. Current robot positions are $(r_1,r_2)$.
 Assume $s_1.x < s_2.x$ and $s_1.y \geq s_2.y$ if not, rotate workspace coordinates $90^{\circ}$.
\Ensure $(e_1, e_2) , (s_1, s_2)$ all at least $\epsilon$ distance from walls.
\State $\epsilon \gets$ robot diameter
\State $\Delta e_y = e_2.y - e_1.y$
\State $\Delta e_x = e_2.x - e_1.x$
%\If {$\Delta e_y \neq 0$}
%\State $sign \gets \frac{|\Delta e_y|}{\Delta e_y}$
%\Else
%\State $sign \gets 0$
%\EndIf
%First Move
\State Move $\left(-r_1.x , -\min \left(r_2.y , r_1.y+ \Delta e_y \right)+\epsilon \right)$ \Comment{Touch left wall}
%Second Move
%(-\min(|-r_2.x+ \Delta e_x|, r_2.x),
\State Move $ (r_2.x, |\sign(\Delta e_y)| \, \min(r_1.y - r_2.y + \Delta e_y-\epsilon,L-r_2.y-\epsilon)+ \epsilon)$ \Comment{Adjust  $y$}
%Third Move
\State Move $(\max(\epsilon, -\sign(r_2.y - r_1.y)\Delta e_x), -\min(r_1.y , r_2.y))$ \Comment{Touch bottom wall}
%Fourth Move
\State Move $(\sign(r_2.y - r_1.y)\Delta e_x , (|\sign(\Delta e_y)|-1) |r_2.y - r_1.y|)$ \Comment{Adjust $x$}
\If {$\Delta e_y \neq r_2.y - r_1.y$}
%Fifth Move
%TODO This step is not correct.
\State Move $(L- \max(r_1.x, r_2.x) , \epsilon)$ \Comment{Touch right wall}
%Sixth Move
\State Move $(0, -\sign(\Delta e_x)\Delta e_y -\max(r_1.y, r_2.y)- \min(r_1.y, r_2.y)$ \Comment{Adjust $y$}
\EndIf
%Seventh Move
\State Move by $(e_2.x - r_2.x, e_2.y - r_2.y)$ \Comment{Go to goal}
\end{algorithmic}
\end{algorithm}
